National Longitudinal Evidence for Growth in Subjective Wellbeing from Spiritual Beliefs

### 2020-11-26 Supplement_And_Analysis.Rmd
few fixes
rewrote supplement

---
title             : "Supplement:Analysis for Longitudinal Evidence for Growth in Subjective Wellbeing from Spiritual Belief"
shorttitle        : "Supplement 1"
author: 
  - name          : "Benjamin Highland"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "520 Power Plant Circle Apartment 105 Winston-Salem, NC 27101 USA"
    email         : "benhighland6420@gmail.com"
  - name          : "Everett L. Worthington"
    affiliation   : "2"
    institution   : "Department of Psychology Virginia Commonwealth University"
  - name          : "Don E Davis"
    affiliation   : "3"
    institution   : "Department of Psychology Georgia State University"
  - name          : "Chris G. Sibley"
    affiliation   : "4"
    institution   : "School of Psychology University of Auckland"
  - name          : "Joseph A. Bulbulia"
    affiliation   : "5"
    institution   : "School of Psychology Victoria University"
authornote: |
  1 Wake Forest School of Medicine 2 Department of Psychology Virginia Commonwealth University 3 Department of Psychology Georgia State University 4 School of Psychology University of Auckland 5 School of Psychology Victoria University

  <!-- https://tinyurl.com/ybremelq -->
# keywords          : "keywords"
wordcount         : no
bibliography      : ["b-references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
# documentclass     : "apa6"
classoption       : "man"
output            : "html_document" 
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
#library("papaja")
#r_refs("r-references.bib")
```

@@

This supplement offers a step-by-step description our full analysis of the "Longitudinal Evidence for Growth in Subjective Wellbeing from Spiritual Belief."  This analysis can be found on github [https://github.com/jacopastorius/nzavs_spiritbeliefWellbeing](https://github.com/jacopastorius/nzavs_spiritbeliefWellbeing)

First we load our functions. We use functions to automate repetative coding, to avoid errors from repetation, and to make the data analysis easier to follow.

```{r load_libraries, echo=TRUE, include=FALSE}
source("libs.R")
source("funs.R")
```

We need to give credit to those who have developed the packages we are using. This script returns a bibliography file.
```{r, sessioninfo, include=FALSE}
#  custom function for citing packages. 
cite_packages()
```

This reads the data into my machine. Note that the NZAVS is an ongoing longitudinal study. Our ethics requires us to protect the privacy of our participants. For this reason, we cannot post the raw data. However, we provide data to qualified researchers for the purposes of replication. Additionally we are interested in collaboration with qualified researchers. For data requests, please contact Joseph Bulbulia [joseph.bulbulia@gmail.com](joseph.bulbulia@gmail.com), Chris G. Sibley  [c.sibley@auckland.ac.nz](c.sibley@auckland.ac.nz).

```{r, read_data, cache=TRUE}
# custom function for my machine to read data. you'll need to read data to your machine. tweak function in the `funs.R' script for your machine
d<-data_read_mymachine()

@@

This code will create a demographic table.

@@

This table shows the  belief responses over the 9 years.  
```{r, include=FALSE}
beliefs_table(d_3)
```

We wrote the publication in \LaTeX. For the data table in LaTeX run this code.

```{r latex_data_table, include=FALSE}
# you might need to install the dev verion.//remotes::install_github("tysonstanley/furniture")
#demographic_table_latex(d_3)
```

We select a minimum of three waves because we want to recover a longitudinal sample in which a slope for the effect of Beliefs might change (or not). However, we want to consider whether our selection of three waves is sensitive to that number. So we also prepare 4-7 years of data frames for separate analyses, and we check the n's for each of those datasets.

```{r, include = FALSE}
# create datasets to test robustness to years
d_4<-data_clean_spirit_wellbeing(d,4) # 4 years
d_5<-data_clean_spirit_wellbeing(d,5) # 5 years
d_6<-data_clean_spirit_wellbeing(d,6) # 6 years
d_7<-data_clean_spirit_wellbeing(d,7) # 7 years
d_8<-data_clean_spirit_wellbeing(d,8) # 7 years

@@

In the first set of analyses we included covariates to adjust for potential confounding effects from usual suspect factors associated with our two measures of wellbeing that might explain the relationship of spirit-beliefs to well-being. We initially included a test of Variance Inflation Factors to rule out mult-collinearity. However, upon further reading, we became aware of at least three other types of confounding that might arise from longitudinal analysis (see: [@mcelreath2020statistical]). We report our controlled analysis below, which does not change the inference in this study. However, because adding covariates can *open* a relationship between beliefs and wellbeing we have switched to reporting the simple time series model in the main text. Additionally, we are careful to state that the present student cannot infer a causal relationshp between spirit beliefs and wellbeing. Rather, we can only infer how the expected relationship between beliefs and wellbeing changes between those who differ in such beliefs over time. A causal understanding inference requires testing specific mechanistic theories. This is a matter for future research.

First we use a maximum Likelihood estimator. Such an estimator is faster than our preferred bayesian estimator.  We consider whether the slope of beliefs changes as a function of time, or equivalently whether the time effect differs depending on one's leve of belief. We also create a table for the two analyses (personal wellbeing and lifesatisfaction.)

```{r robust to years and beliefs only, cache = TRUE, dependson=d_3,cache =TRUE}
library(lme4)
pwi_formula_simple <- 'PWI ~  years * Beliefs + (1+ Beliefs|Id)'
ls_formula_simple <- 'LIFESAT ~  years * Beliefs + (1+ Beliefs|Id)'
mod.1aa <-lmer(pwi_formula_simple, data= d_3)
mod.2aa <-lmer(ls_formula_simple, data= d_3)

@@

We next produce a predictive graph for beliefs by time on PWI. Note, again, the function for producing this graph is in the "funs.R" file.

```{r prediction plot of ML longitudinal model, cache =TRUE}
```{r bayesian_pwi_predict, cache =TRUE}
graph_predictions(mod.1aa, "Predicted Values of Personal Wellbeing")
```

We next produce a predictive graph for beliefs by time on life satisfaction. 
```{r bayesian_lifesat_predict, cache =TRUE}
graph_predictions(mod.2aa, "Predicted Values of Life Satisfaction")
```

For both outcomes, the slope for beliefs in a spirit or life fource predicts a greater increase in  wellbeing.

Next we attempt a model with that is adjusted by "usual suspect" demographic inicators. We find that the relationship is a little stronger. Because these strengthened relationship might owe to confounding, we are wearing of the adjusted model.  The idea that skeptics are not increasing in well-being, whereas other groups are increasing strikes us as improbable. We do not find evidence for a such a stall in the bayesian analysis; nor do we find any such evidence for a stall in the multiply-imputed dataset analysis (which handles missing responses).  Ultimately we favour, and report the bayesian analysis because bayesian data analysis recovers an explicit estimate of posterior uncertianty.

```{r lmer_model_with_full_covariates, cache =TRUE}
library(lme4)
pwi_formula <- 'PWI ~  years * Beliefs +  Age_in_Decades_C + Deprivation_S + Education_S  + Employed + Ethnic_Categories + Male  + Has_Partner  + Political_Conservativism_S + Urban + (1 + Beliefs|Id)'
ls_formula <- 'LIFESAT ~  years * Beliefs +  Age_in_Decades_C + Deprivation_S + Education_S  + Employed + Ethnic_Categories + Male  + Has_Partner  + Political_Conservativism_S + Urban + (1 + Beliefs|Id)'

@@

Compare complex with simple models, and we find the inference doesn't change. 
```{r}
# identical result
sjPlot::tab_model(mod.1,mod.1aa)
sjPlot::tab_model(mod.2,mod.2aa)
```

Next we check for multi-collinearity.

```{r check collinearity, cache =TRUE}
# Note High VIFS from 
performance::check_collinearity(mod.1)
performance::check_collinearity(mod.2)
```

We observe that VIFs are high, for the years x beliefs interaction. However, this is likely to be an artifact of how we code the variables (we do not center by default), but to make sure let's center the years indicator and run the models again. We we do this we find that collinearity disappears, and without any effect on inference.

```{r lmer model centering years}
library(lme4)
pwi_formulaC <- 'PWI ~  yearsC * Beliefs +  Age_in_Decades_C + Deprivation_S + Education_S  + Employed + Ethnic_Categories + Male  + Has_Partner  + Political_Conservativism_S + Urban + (1 |Id)'

ls_formulaC <- 'LIFESAT ~  yearsC * Beliefs +  Age_in_Decades_C + Deprivation_S + Education_S  + Employed + Ethnic_Categories + Male  + Has_Partner  + Political_Conservativism_S + Urban + (1 |Id)'

@@

summary(mod.1a)
summary(mod.2a)
# identical result
performance::check_collinearity(mod.1a)
performance::check_collinearity(mod.2a)
```

Here's the table, again, inference is robust.

@@

Her are thecoefficient plots for the full models
```{r coefficientplots_LMER_FULL, cache=TRUE}
plot_models(mod.2, mod.1) + ylim(-.4, .75) + theme_abyss() + scale_colour_brewer(labels = c("Personal Wellbeing", "Life Satisfaction")) # Plots
```

Here is the code for the coefficient plots of our preferred simple model
```{r cache =TRUE}
g1<-plot(parameters::model_parameters(mod.1aa))+ ggtitle("Personal Wellbeing Coefficients")
g2<-plot(parameters::model_parameters(mod.2aa))+ ggtitle("Life Satisfaction Coefficients")
```

@@

Next, recall we want to check whether inference is  robustness to the researcher degree of freedom in selecting the number of waves.
```{r run all the models with different ns, cache = TRUE}
mod.1_3 <-lmer (pwi_formula_simple, data= d_4)
mod.2_3 <-lmer (ls_formula_simple, data= d_4)

@@

Let's plot all the models.  We we do so, we find that uncertainty increases, but that the inference remains robust.

Let's graph the extremes of the degree-of-freedom choice to be certain that the inference remains satisfying to the eye (and it is.)

```{r extremes_of_researcher_degree_of_freedom_test, cache =TRUE}
library(parameters)
g1a_3<-plot(parameters::model_parameters(mod.1_3))+ ggtitle("Personal Wellbeing Coefficients: 8 Waves")
g2a_3<-plot(parameters::model_parameters(mod.2_3))+ ggtitle("Life Satisfaction Coefficients: 8 Waves")

g1a_8<-plot(parameters::model_parameters(mod.1_8))+ ggtitle("Personal Wellbeing Coefficients: 8 Waves")
g2a_8<-plot(parameters::model_parameters(mod.2_8))+ ggtitle("Life Satisfaction Coefficients: 8 Waves")
```

Estimates robust at 8 waves
```{r USE_2_coefficientplots_LMER_FULL,cache =TRUE}
library(patchwork)
g1a_3/g1a_8
```

Finally we model the relationshp bewteen having a partner and deprivation, to address the reviewer's worry that we might have introduced correlation of between deprivation and having a partner that could spoil inference. We indeed find such a relationship. This is an important potential confound. Again we emphasise that in a longitudinal setting the reviewers concern is even greater, as we might open a path to the outcome variable by including the demographic indicator.

@@

# Bayesian estimation

Bayesian inference is time intensive. However, a bayesian model will pool information to produce a better, and natural estimate of uncertainty. The posterior distributions of the model describe conditional probabilities. So where we can we want to use bayesian models. Here we use the default weakly informative priors that the BRMS package produces. It is generally better to use informative priors. However, the size or our dataset does not make a difference. We report the priors at the end.

```{r setmodelformula}
forumula_bayesPWI <- bf(PWI ~ years * Beliefs + (1 + Beliefs|Id))
forumula_bayesLIFESAT <- bf(LIFESAT ~  years * Beliefs + (1 + Beliefs|Id))
```
```{r pwi_bayesmodel}
# PWI
mod.1_b <-brm(forumula_bayesPWI, data= d_3, file ="mod.1_b.rds") # note these are very large files
summary(mod.1_b)
#saveRDS(mod.1_b,"backup_mod.1_b") this is excessive, but we save another backup to avoid running this model again in case of SNAFU.
rm(mod.1_b)
```

```{r lsbayesmodel}
# LIFE SAT
mod.2_b <-brm (forumula_bayesLIFESAT, data= d_3,cores = 8, file ="mod.2_b.rds") # note these are very large files
# saveRDS(mod.2_b,"backup_mod.2_b") this is excessive, but we save another backup to avoid running this model again in case of SNAFU.
```

Tables:
```{r lifesat bayesian model table:pwi, cache =TRUE}
tab_model(mod.1_b)
```

```{r lifesat bayesian model table:ls lifesat, cache =TRUE}
tab_model(mod.2_b)
```

Here we extract the information about the priors

```{r stancode for models}
#model equation
stancode(mod.1_b)
stancode(mod.2_b)
```

```{r prior summaries}
brms::prior_summary(mod.1_b)
brms::prior_summary(mod.2_b)
```

```{r bayesian_model_table, cache=TRUE}
#tab_model(mod.1_b, show.ci= .89) # we really should use something other than 95 but the team consensus is to use 95
tab_model(mod.1_b, mod.1_b show.ci= .95) # no difference to inference
```

```{r mixing of sd_pwi}
plot(mod.1_b, pars ="b_") 
```

```{r mixing of ls_pwi}
plot(mod.1_b, pars ="sd_") # requires stronger priors
```

# ```{r posterior predictive checks bayes:pwi, cache=TRUE}
# # cannot run on my laptop
# # pp_1<-rstanarm::pp_check(mod.1_b) + ylab("Personal Wellbeing")
# # pp_2<-rstanarm::pp_check(mod.2_b) + ylab("Life Satisfaction")
# # 
# # # plot
# pp_1 + pp_2
#```

Here are the prediction plots that we use in the main publication text.

Predictions and plots of them. First we get the predictions.

```{r bayesian_pwi_predicted, cache=TRUE, dependson= mod.1_b}
graph_predictions(mod.1_b, "Predicted Values of Personal Wellbeing")
```
```{r bayesian_lifesat_predicted, cache=TRUE, dependson= mod.2_b}
graph_predictions(mod.2_b, "Predicted Values of Life Satisfaction")
```

## Imputed Analysis

Finally we run the data analysis to test whether our results are robust to missing data. Missing responses can be an important source of bias in a longitudinal study. Again we write some functions to deal with all the work. I have saved the imputed files and will read them back to for analysis.

```{r echo = TRUE,  cache = TRUE, dependson = "dfben"}
# Imputed data
# imputed3<-amelia_imputation_clean(d_3)
# saveRDS(imputed3,"imputed3")
imputed3<-readRDS(imputed3)
```

We rerun the imputed dat analysis.

```{r cache = TRUE, imputedsimplePWI}
# run models iterating over imputed data
mod_imputed_pwi<-loop_lmer_model(pwi_formula_simple,imputed3)
```

```{r imputed pwi table, cache=TRUE}
tab_mods_ls<-loop_lmer_model_tab(mod_imputed_ls)
create_hux_table(tab_mods_ls)
```

We graph the coefficients.
```{r coef_PWI_imputed coefplot, cache =TRUE}
# coeff plot
imp_pwi <- plot(tab_mods_pwi) + ggtitle("Personal Wellbeing Coefficients (Imputed)") # pick your  own title
imp_pwi
```

We create a graph for the predictions.
```{r r coef_PWI_imputed coefplot, cache=TRUE}
# function for prediction
graph_predictions_imputed(mod_imputed_pwi,"Predicted Values of Personal Wellbeing (imputed)")
```

@@

## Life Satisfaction Imputation Model

The following code returns the analysis for the life statisfaction model (again the functions are in the "funs.R" file).

@@

```{r table_ls_imputed,cache=TRUE}
# table
tab_mods_ls<-loop_lmer_model_tab(mod_imputed_ls)
create_hux_table(tab_mods_ls)
```

@@

```{r coef_PWI_imputed, cache = TRUE}
# coeff plot
imp_ls<-plot(tab_mods_ls) + ggtitle("Life Satisfaction Coefficients (Imputed)") 
imp_ls
```

>>>>

---
title             : "Supplement: Analysis for Longitudinal Evidence for Growth in Subjective well-being from Spiritual Belief"
author: "contact: joseph.bulbulia@gmail.com on behalf of Highland et al."
date: "27/11/2020"
authornote: |
  <!-- https://tinyurl.com/ybremelq -->
# keywords          : "keywords"
wordcount         : no
bibliography      : ["b-references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
# documentclass     : "apa6"
classoption       : "man"
output            : "html_document" 
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
# library("papaja")
#r_refs("r-references.bib")
```

@@

This supplement offers a step-by-step analysis for the "Longitudinal Evidence for Growth in Subjective Well-being from Spiritual Beliefs." The functions and other materials for this analysis can be found at the following GitHub link: [https://github.com/jacopastorius/nzavs_spiritbeliefwell-being](https://github.com/jacopastorius/nzavs_spiritbeliefwell-being)

First, we load our functions.

```{r load_libraries, echo=TRUE, include=TRUE}
source("libs.R")
source("funs.R")
```

We use functions to automate repetition in coding, to avoid errors from such repetition, and to make the data analysis easier to follow.

This function reads our data into my machine. 
```{r, read_data, cache=TRUE}
# custom function for my machine to read data. you'll need to read data to your machine. tweak function in the `funs.R` script for your machine
d<-data_read_mymachine()

@@

## Sample demographic table

@@

```{r, include=FALSE}
beliefs_table(d_3)
```

We wrote the publication in $\LaTeX$, which requires converting the tables to .tex files; this conversion used a custom function we wrote in the `funs.R` file.

```{r latex_data_table, include=TRUE}
# you might need to install the dev verion.//remotes::install_github("tysonstanley/furniture")
#demographic_table_latex(d_3) # note the furniture package conflicts with the table1 package. 
```

## How many responses waves to select?

We selected a minimum of three waves because we want to recover a longitudinal sample in which a slope for the effect of the years $\times$ beliefs effect might be recovered. However, we needed to consider whether our selection of three waves was sensitive to that degree of researcher freedom. So we also prepared 4-8 years of data frames for separate analyses.

```{r, include = TRUE}
# create datasets to test robustness to years
d_4<-data_clean_spirit_wellbeing(d,4) # 4 years
d_5<-data_clean_spirit_wellbeing(d,5) # 5 years
d_6<-data_clean_spirit_wellbeing(d,6) # 6 years
d_7<-data_clean_spirit_wellbeing(d,7) # 7 years
d_8<-data_clean_spirit_wellbeing(d,8) # 7 years

@@

The analyses for these waves is described below. These analyses reveal that the inference of the paper is not sensitive to the threshold number of waves for participant selection. Which is to say, we recover the same inference with an 8-wave selection criterion.

## Covariates

In the the reported analyses, we included covariates to adjust for potential confounding effects from usual suspect factors associated with our two measures of well-being (Personal Well-Being, Life Satisfaction). Where $Z$ is a third variable, $X$ is the predictor, and $Y$ is the outcome, there are at least four types of confounding that might arise: 
1. $Z$ causes $X$ and  $Z$ causes $Y$: $Z \leftarrow Z \rightarrow Y$
2. $X$ causes $Z$ and  $Z$ causes $Y$: $X \rightarrow Z \rightarrow Y$
3. $X$ causes $Z$ and $Y$ causes $Z$: $X \rightarrow Z; Y \rightarrow Z$
4. $X$ causes $Z$; $Z$ causes $D$; and we condition on D (the descendent of $Z$: $Y$ causes $Z$: $Z$ \rightarrow D; $X \rightarrow Z; Y \rightarrow Z$ (see: [@mcelreath2020statistical], Chapter 6).

We report our usual-suspect analysis below, which does not change the inference in this study. However, the third and fourth type of confounding illustrate how conditioning on a covariates may confound by *opening* a relationship between the $X$ and $Y$ indicators in a regression analysis. Such confounding is particularly worrying in the case of longitudinal research.  The second type of counfounding can mask a relationship. We do not wish to exclude third variable explanations of the kind that arise from the first type of confounding. Indeed, we are careful to state in the main text that our study cannot infer a causal relationship between spirit beliefs and well-being. Rather, the contribution of the study comes as an inference about how the expected relationship between beliefs and well-being changes between those who differ in such beliefs over time. By contrast, a causal inference requires testing specific mechanistic theory. This is a matter for future research. Finally, the usual suspect analysis that we report below does not affect inference. We find that the relationships that we observe in the simple analysis also hold there.

## Bayesian data analysis

Bayesian inference require substantial computational resources. The cost of time is repaid with a model that efficiently pools information to produce a more accurate, and transparent, estimate of uncertainty. The posterior distributions of the model describe intuitive conditional probabilities.

Below we write the model in a notation that explicitly states the the default weakly informative priors that the BRMS package produces. And indicated in the model we include group-level indicator for variation in subject-level intercepts $\alpha_j$, a group-level indicator for subject-level time (or "growth")  slope  $\Beta_j$, and we model the co-variance of $\alpha_j$, Below we write the model in a notation that explicitly states the the default weakly informative priors that the BRMS package produces.  And indicated in the model we include group-level indicator for variation in subject-level intercepts $\alpha_j$, a group-level indicator for subject-level time (or "growth")  slope  $\Beta_j$, and we model the co-variance of $\alpha_j$ and  $\Beta_j$ which for computational reasons is estimate as a correlation matrix of (${\bf S}$} and ${\bf R}$).

### Priors for the Personal Well-Being model

First we report priors for the Personal Well-Being model. Note that we model the correlation of the id-level intercept and id-level year-slope.

### Priors for the Life Satisfaction model:

Model priors for the  Life Satisfaction model. Here, again, we model the correlation of the id-level intercept and id-level year-slope.

\[
\begin{aligned}
\operatorname{Life Satisfaction}_{ij} \sim \text{Normal}(\mu_{ij},\sigma^2) \nonumber\\
\mu_{ij}\sim \boldsymbol{\alpha +\beta} \nonumber\\
\boldsymbol{\alpha} \sim \alpha_{0} +\alpha_j \nonumber\\
\alpha_{0} \sim \text{StudentT}(3,5.5, 2.5)\nonumber\\
\alpha_j \sim \text{StudentT}+(3,0,2.5)\nonumber\\
\boldsymbol{\beta}\sim \boldsymbol{b} + \beta_j \nonumber\\
\boldsymbol{b}\sim\text{Normal}(0,2.5) \nonumber\\
\begin{bmatrix}
\alpha_j \\
\beta_j
\end{bmatrix}
\sim 
\text{MVNormal}
\begin{pmatrix}
\begin{bmatrix}
0\\
0
\end{bmatrix}
,\boldsymbol{S} \nonumber
\end{pmatrix}\\
\boldsymbol{S} = 
\begin{pmatrix}
\sigma_{\alpha_j} & 0 \\
0 & \sigma_{\beta_j} \nonumber
\end{pmatrix} \boldsymbol{R} \begin{pmatrix}
\sigma_{\alpha_j} & 0 \\
0 & \sigma_{\beta_j} \nonumber
\end{pmatrix}\\
\sigma_{\alpha_j} \sim 
\text{HalfCauchy}(0,1) \nonumber\\
\sigma_{\beta_j} \sim 
\text{HalfCauchy}(0,1) \nonumber\\
\boldsymbol{R}\sim \text{LKJcorr}(1) \nonumber\\
\sigma  \sim  \operatorname{HalfStudent T}^{+}(3, 0, 2.5) 
\end{aligned}
\]

### Theoretical model

The deterministic part of both the Personal Well-Being and the Life Satisfaction models is as follows.

```{r setmodelformula}
forumula_bayesPWI <- bf(PWI ~ yearsC * Beliefs + (1 + yearsC|Id))
forumula_bayesLIFESAT <- bf(LIFESAT ~  yearsC * Beliefs + (1 + yearsC|Id))
```
```{r pwi_bayesmodel}
# PWI
mod.1_b <-brm(forumula_bayesPWI, data= d_3, file ="mod.1_b.rds", cores = 8) # note these are very large files
```

```{r lsbayesmodel}
# LIFE SAT
mod.2_b <-brm (forumula_bayesLIFESAT, data= d_3,cores = 8, file ="mod.2_b.rds") 
# note these are very large files
```

<!-- Tables: -->
<!-- # ```{r pwi_Bayesian_model_table_pwi, cache =TRUE} -->
<!-- # tab_model(mod.1_b) -->
<!-- # ``` -->
<!-- # -->
<!-- # ```{r lifesat_Bayesian_model_table_ls lifesat, cache =TRUE} -->
<!-- # tab_model(mod.2_b) -->
<!-- # ``` -->

```{r texreg_bayes_table:pwi,cache=TRUE, include=FALSE}
#Results for publication.
table_model_latex_pwi(mod.1_b)
```

```{r texreg_bayes_table:ls, cache=TRUE, include=FALSE}
table_model_latex_ls(mod.2_b)
```

<!-- Coefficient plots used in manuscript-->
<!-- ```{r coefficientplotsbayes, cache=TRUE} -->
<!-- plot_coefs(mod.1_b,mod.2_b) -->
<!-- ``` -->

```{r stancode for models, include=FALSE}
#model equation
#Here we extract our priors for reporting

stancode(mod.1_b)
stancode(mod.2_b)
```

```{r prior summaries, include =FALSE}
#this does the same as the stancode
brms::prior_summary(mod.1_b)
brms::prior_summary(mod.2_b)
```

```{r Bayesian_model_table, cache=TRUE}
#tab_model(mod.1_b, show.ci= .89) # we really should use something other than 95 but the team consensus is to use 95
tab_model(mod.1_b, mod.2_b, show.ci= .95) # no difference to inference
```
<!-- ```{r Bayesian_model_table, cache=TRUE, include =FALSE} -->
<!-- table_model_latex_dual() -->
<!-- ``` -->

We inspected the posterior distributions, which look good.

```{r mixing_pwi, cache=TRUE}
plot(mod.1_b, pars ="b_") 
```
```{r mixing_ls, cache=TRUE}
plot(mod.2_b, pars ="b_") 
```

As do the random components

```{r mixing_sd_pwi}
plot(mod.1_b, pars ="sd_") # requires stronger priors
```

```{r mixing_ls_pwi}
plot(mod.2_b, pars ="sd_") # requires stronger priors
```

The posterior prediction plots look reasonable.

```{r posteriorpredictivechecks:pwi, cache=TRUE}
# cannot run on my laptop
pp_1<-rstanarm::pp_check(mod.1_b) + ylab("Personal well-being")
pp_2<-rstanarm::pp_check(mod.2_b) + ylab("Life Satisfaction") ## life sat suggests an ordinal model
pp_1 + pp_2
```

### Predictions graphs

Personal Well being predictions.

Life Satisfaction predictions.

Next we graph the expected effects.

```{r Bayesian_pwi_predicted, cache=TRUE, dependson= mod.1_b}
graph_predictions(mod.1_b, "Predicted Values of Personal Well-Being")
```
```{r Bayesian_lifesat_predicted, cache=TRUE, dependson= mod.2_b}
graph_predictions(mod.2_b, "Predicted Values of Life Satisfaction")
```

## ML estimation

Many researchers have yet to familiarise themselves with Bayesian estimation. We therefore report results from models using a Maximum Likelihood estimator. We do so to avoid the perception of chicanery, as our experience is that such a choice does not affect inference for a dataset such as ours.  ML estimatation is also faster than our preferred Bayesian estimator.

Results are virtually identical to the Bayesian model.

```{r robust to years and beliefs only, cache = TRUE, dependson=d_3,cache =TRUE}
library(lme4)
pwi_formula_simple <- 'PWI ~  yearsC * Beliefs + (1+ yearsC|Id)'
ls_formula_simple <- 'LIFESAT ~  yearsC * Beliefs + (1+ yearsC|Id)'
mod.1aa <-lmer(pwi_formula_simple, data= d_3)
mod.2aa <-lmer(ls_formula_simple, data= d_3)

@@

Here is a predictive graph for years $\times$ beliefs interaction on PWI. Rhe function for producing this graph is in the `funs.R` file.

```{r Bayesian_pwi_predict, cache =TRUE}
graph_predictions(mod.1aa, "Predicted Values of Personal well-being")
```

Here is a predictive graph for the years $\times$ beliefs interaction on life satisfaction.

```{r Bayesian_lifesat_predict, cache =TRUE}
graph_predictions(mod.2aa, "Predicted Values of Life Satisfaction")
```

## Adjusting for Demographics.

Next we attempted a model with that is adjusted by "usual suspect" demographic indicator. The inference remains unchanged.

```{r lmer_model_with_full_covariates, cache =TRUE}
library(lme4)
pwi_formula <- 'PWI ~  yearsC * Beliefs +  Age_in_Decades_C + Deprivation_S + Education_S  + Employed + Ethnic_Categories + Male  + Has_Partner  + Political_Conservativism_S + Urban + (1 + yearsC|Id)'
ls_formula <- 'LIFESAT ~  yearsC * Beliefs +  Age_in_Decades_C + Deprivation_S + Education_S  + Employed + Ethnic_Categories + Male  + Has_Partner  + Political_Conservativism_S + Urban + (1 + yearsC|Id)'

@@

We compared complex with simple models, and we found the inference does not change. 
```{r}
# identical result
sjPlot::tab_model(mod.1,mod.1aa)
sjPlot::tab_model(mod.2,mod.2aa)
```

Next we checked for multi-collinearity. Multi-collinearity is low. (Note that we centre the "years" variable, had we not done this, mult-collinearity would have been high).

```{r check collinearity, cache =TRUE}
# Low VIFS from 
performance::check_collinearity(mod.1)
performance::check_collinearity(mod.2)
```

```{r lmer model centering years}
library(lme4)
pwi_formulaC <- 'PWI ~  yearsC * Beliefs +  Age_in_Decades_C + Deprivation_S + Education_S  + Employed + Ethnic_Categories + Male  + Has_Partner  + Political_Conservativism_S + Urban + (1+ yearsC|Id)'

ls_formulaC <- 'LIFESAT ~  yearsC * Beliefs +  Age_in_Decades_C + Deprivation_S + Education_S  + Employed + Ethnic_Categories + Male  + Has_Partner  + Political_Conservativism_S + Urban + (1+ yearsC|Id)'

@@

# identical result
performance::check_collinearity(mod.1a)
performance::check_collinearity(mod.2a)
```

Here's the table, again, we find inference is robust.

@@

<!-- Her are the coefficient plots for the full models -->
<!-- ```{r coefficientplots_LMER_FULL, cache=TRUE} -->
<!-- plot_models(mod.2, mod.1) + ylim(-.4, .75) + theme_plain()  # Plots -->
<!-- ``` -->

Here is the code for the coefficient plots of our preferred simple model.

```{r cache =TRUE}
g1<-plot(parameters::model_parameters(mod.1aa))+ ggtitle("Personal well-being Coefficients")
g2<-plot(parameters::model_parameters(mod.2aa))+ ggtitle("Life Satisfaction Coefficients")
```

@@

Next, recall we want to check whether inference is robustness to the researcher degree of freedom in selecting how many number of waves a participant must have responded to for inclusion.

```{r run all the models with different ns, cache = TRUE}
mod.1_3 <-lmer (pwi_formula_simple, data= d_4)
mod.2_3 <-lmer (ls_formula_simple, data= d_4)

@@

Let's plot all the models. We find that uncertainty increases, but that, again, the inference remains robust.

```{r extremes_of_researcher_degree_of_freedom_test, cache =TRUE}
library(parameters)
g1a_3<-plot(parameters::model_parameters(mod.1_3))+ ggtitle("Personal well-being Coefficients: 8 Waves")
g2a_3<-plot(parameters::model_parameters(mod.2_3))+ ggtitle("Life Satisfaction Coefficients: 8 Waves")

g1a_8<-plot(parameters::model_parameters(mod.1_8))+ ggtitle("Personal well-being Coefficients: 8 Waves")
g2a_8<-plot(parameters::model_parameters(mod.2_8))+ ggtitle("Life Satisfaction Coefficients: 8 Waves")
```

```{r USE_2_coefficientplots_LMER_FULL_pwi,cache =TRUE}
library(patchwork)
g1a_3/g1a_8 +   plot_annotation(title = "Compare 3 X response model (top) with 8 X response model (bottom): Personal Well-Being")
```

```{r USE_2_coefficientplots_LMER_FULL_ls,cache =TRUE}
library(patchwork)
g2a_3/g2a_8 +   plot_annotation(title = "Compare 3 X response model (top) with 8 X response model (bottom) : Life Satisfaction")
```

Next, we modelled the relationship between having a partner and deprivation. We did so to address a reviewer's worry that we might have introduced correlation of between deprivation and having a partner that could spoil inference. We indeed found such a relationship. This is an important potential confound, and another reason to favour the simple longitudinal model that we used.

@@

# Missing Data Analysis

Finally, we ran our analyses on 10 multiply imputed datasets. Missing responses can be an important source of bias in a longitudinal study. Again we wrote some functions to deal with all the work, you can find these in `funs.R` file.

```{r echo = TRUE,  cache = TRUE}
# Imputed data
imputed3<-amelia_imputation_clean(d_3)
# saveRDS(imputed3,"imputed3")
# imputed3<-readRDS(imputed3)
```

We reran the imputed data analysis.

```{r cache = TRUE, imputedsimplePWI, cache=TRUE}
# run models iterating over imputed data
mod_imputed_pwi<-loop_lmer_model(pwi_formula_simple, imputed3)
```

```{r imputed pwi table, cache=TRUE}
tab_mods_pwi<-loop_lmer_model_tab(mod_imputed_pwi)
#create_hux_table(tab_mods_pwi)
```

Next a graph of the coefficients.

```{r coef_PWI_imputed_coefplot, cache =TRUE}
# coeff plot
imp_pwi <- plot(tab_mods_pwi) + ggtitle("Personal well-being Coefficients (Imputed)") # pick your  own title
```

@@

The following code returns the analysis for the life satisfaction model (again the functions are in the `funs.R` file).

@@

```{r table_ls_imputed,cache=TRUE}
# table
tab_mods_ls<-loop_lmer_model_tab(mod_imputed_ls)
#create_hux_table(tab_mods_ls)
```

@@

```{r coef_imputed_ls, cache = TRUE}
# coeff plot
imp_ls<-plot(tab_mods_ls) + ggtitle("Life Satisfaction Coefficients (Imputed)") 
```
We found that the model is robust to multiple imputations over the missing data, though with greater uncertainty for the Personal Well-being effect.

Here are the prediction graphs.

```{r  coef_PWI_imputed coefplot, cache=TRUE}
# function for prediction
graph_predictions_imputed(mod_imputed_pwi,"Predicted Values of Personal well-being (imputed)")
```
